---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ecdm

<!-- badges: start -->
[![R build status](https://github.com/tmsalab/ecdm/workflows/R-CMD-check/badge.svg)](https://github.com/tmsalab/ecdm)
[![Package-License](http://img.shields.io/badge/license-GPL%20(%3E=2)-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-2.0.html)
[![CRAN Version Badge](http://www.r-pkg.org/badges/version/edina)](https://cran.r-project.org/package=edina)
[![CRAN Status](https://cranchecks.info/badges/worst/edina)](https://cran.r-project.org/web/checks/check_results_edina.html)
[![RStudio CRAN Mirror's Monthly Downloads](http://cranlogs.r-pkg.org/badges/edina?color=brightgreen)](http://www.r-pkg.org/pkg/edina)
[![RStudio CRAN Mirror's Total Downloads](http://cranlogs.r-pkg.org/badges/grand-total/edina?color=brightgreen)](http://www.r-pkg.org/pkg/edina)
<!-- badges: end -->

The goal of ecdm is to provide a modeling framework for exploratory cognitive
diagnostic models and classical cognitive diagnostic models.

## Installation

The `ecdm` package is currently only available via GitHub. To install `ecdm`,
your computer will need to have a compiler. The following guides are avaliable:

- [Windows: Rtools](http://thecoatlessprofessor.com/programming/installing-rtools-for-compiled-code-via-rcpp/)
- [macOS: Rtools](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-macos/)

From there, please use `devtools` to retrieve the latest development version.

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tmsalab/ecdm")
```


## Usage

Load the `ecdm` package into _R_: 

```r
library(ecdm)
```

Exploratory CDM models can be estimated with:

```r
edina_model = edina(<data>, <k>)
errum_model = errum(<data>, <k>, ... )
```

Classical CDMs can be estimated using:

```r
dina_model = dina(<data>, <q>)
rrum_model = rrum(<data>, <q>)
```

## Details

The `ecdm` package is designed to act more as a "virtual" package. The main
functionalities of `ecdm` are split across multiple packages. The rationale
for this is many areas of psychometrics have overlap in terms of computational
code used. By dividing the underlying source of the `ecdm` package, we are
enabling fellow psychometricians to be able to incorporate established 
routines into their own code. In addition, we are lowering the amount of
redundancies, or copy and pasted code, within the CDM framework we are building.

Specifically, the `ecdm` package imports estimation routines from:

- `dina`: Estimating the Deterministic Input, Noisy "And" Gate ('DINA') cognitive
   diagnostic model parameters using a Gibbs sampler.
- `edina`: Estimating the Exploratory Deterministic Input, Noisy "And" Gate 
   ('EDINA') cognitive diagnostic model parameters using a Gibbs sampler.
- `rrum`: Estimating the reduced Reparametrized Unified Model ('rRUM') with a 
  Gibbs sampler.
- `errum`: Estimating the Exploratory reduced Reparametrized Unified Model 
  ('ErRUM') with a Gibbs sampler.

Moreover, we have additional packages that are used within the modeling process:

- `rgen`: Simulate Multivariate Probability Distributions
- `simcdm`: Simulate responses underneath a DINA or rRUM model.
- `shinyecdm`: User Interface for Modeling with Exploratory Models

Lastly, we hace sampled data packages available here: 

- `ecdmdata`: Data package containing psychometric modeling data used in multiple
   packages. 
   
## Authors

James Joseph Balamuta, Steven Andrew Culpepper, and Jeffrey A. Douglas
   
## Citing the `ecdm` package

To ensure future development of the package, please cite `ecdm` package
if used during an analysis or simulation studies. Citation information
for the package may be acquired by using in *R*:

``` r
citation("ecdm")
```
   
## License

GPL (>= 2)
