---
title: "Overview of ecdm"
author: "James Balamuta"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of ecdm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ecdm)
library(ecdmdata)
```


# Introduction to Exploratory Cognitive Diagnostic Models 

Exploratory Cognitive Diagnostic Models (ECDMs) are versions of classicial 
Cognitive Diagnostic Models (CDMs) that do _not_ require the component of 
an expertly crafted _Q_ matrix. This class of models is new to the world of
psychometric models. The goal of the `ecdm` package is to provide an official
implementation of the methodology by the authors that developed it!

Before we continue, please "bookmark" the TMSALab organization on GitHub:

**<https://github.com/tmsalab/>**

The website provides direct access to the developers behind the estimation 
packages being discussed today. In particular, it features the ability to file 
issues or bug reports, ask questions, or stay up-to-date in the latest
breakthroughs.

## Installation

In this section, we show the steps required to _install_ the package locally.
Please note, that the `ecdm` package is currently only available via GitHub as
it is still being developed. As a result, you presently _cannot_ install the
package using `install.packages('ecdm')`. 

Presently, to install the `ecdm` package, your computer will need to have a 
compiler. To assist in setting up the compiler, we've created the following
guides:

- [Windows: Rtools](http://thecoatlessprofessor.com/programming/installing-rtools-for-compiled-code-via-rcpp/)
- [macOS: Rtools](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-macos/)

From there, please use `devtools` to retrieve the latest development version.

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tmsalab/ecdm")
```

## Loading the Package

Accessing the `ecdm` rountines requires loading the package into _R_. Please
load the `ecdm` package by pressing "run"

```{r main-lib}
library(ecdm)
```

The `ecdm` package has an accompanying data package called `ecdmdata`. Please
load this package into _R_ by substituting `ecdm` with `ecdmdata` in the `library()`
call.

```{r secondary-lib-solution}
library( ecdmdata )
```

## Supplementary Data Sets

The `ecdmdata` package presently comes equipped with three different data sets:

- Examination for the Certificate of Proficiency in English (ECPE), Templin, J. and Hoffman, L. (2013). 
    - `items_ecpe`, N = 2922 subject responses to J = 28 items.
    - `qmatrix_ecpe`, J = 28 items and K = 3 traits. 
- Fraction Addition and Subtraction, Tatsuoka, K. K. (1984).
    - `items_fractions`: N = 536 subject responses to J = 20 items.
    - `qmatrix_fractions`: J = 20 items and K = 8 traits. 
- Revised PSVT:R, Culpepper and Balamuta (2013).
    - `items_spatial`: N = 516 subject responses to J = 30 items.
    
We will likely add more data sets as time goes on.

Let's take a look at Fraction Addition and Subtraction data sets. Typing the
name of each data set and running the command will load the data into _R_ if
the `ecdmdata` package is loaded. As these data sets are relatively big, 
let's use the function `head()` to view on the first 6 rows.

```{r data-inc-solution}
head( items_fractions )
head( qmatrix_fractions )
```

Within this guide, we will use the following notation:

- $K$: Number of Traits (columns) in the Q Matrix
- $J$: Number of Items (rows/columns) in the Q Matrix and Item Matrix
- $N$: Number of Subjects (rows) in the Item Matrix

To retrieve this information in _R_, we can use the dimension function,
`dim()`, which lists the size of the data as rows by columns. 

Find the dimensions of the `items_fractions` and `qmatrix_fractions`

```{r dim-info-solution}
dim(items_fractions)
dim(qmatrix_fractions)
```

## Help!

Each function within the package contains a help file that is associated with.
Some of the functions have _worked_ exercises as well. To view this information
type either `?function_name` or `help(function_name)`. Let's verify the
previously acquired numbers for the `items_fractions` data set by checking
the entry in the documentation.

```{r help-info-solution}
?items_fractions
?qmatrix_fractions
```

If you are curious to see how a function performs, you can opt to use 
`example(function_name, package = "ecdm")`. Be aware that some examples
may take considerably longer than the rest.
